---
title: "king_dav_ps04"
author: "Dav King"
date: "3/23/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(socviz)
library(tidycensus)
library(tigris)
library(viridis)
options(tigris_use_cache = T)
```

```{r pop-names}
pop_names <- tribble(
    ~varname, ~clean_name,
    "B01003_001", "pop",
    "B01001B_001", "black",
    "B01001A_001", "white",
    "B01001H_001", "nh_white",
    "B01001I_001", "hispanic",
    "B01001D_001", "asian"
  )
  
pop_names

inc_names <- tribble(
    ~varname, ~clean_name,
    "S1901_C01_012", "median_hh_inc")
```

```{r county-data}
fips_pop <- get_acs(geography = "county", 
                    variables = pop_names$varname, 
                    cache_table = TRUE) %>% 
  mutate(variable = reduce2(pop_names$varname, 
                            pop_names$clean_name, 
                            str_replace, 
                            .init = variable)) %>% 
  select(-moe) %>% 
  pivot_wider(names_from = variable, values_from = estimate) %>% 
  rename(fips = GEOID, name = NAME) 
```

```{r income-data}
fips_inc <- get_acs(geography = "county", 
                    variables = inc_names$varname,  
                    cache_table = TRUE) %>% 
  mutate(variable = str_replace(variable, 
                            inc_names$varname, 
                            inc_names$clean_name)) %>% 
  rename(fips = GEOID, name = NAME) 
```

```{r spatial-data}
fips_map <- get_acs(geography = "county", 
                    variables = "B01001_001", 
                    geometry = TRUE,
                    shift_geo = FALSE,
                    cache_table = TRUE) %>% 
  select(GEOID, NAME, geometry) %>% 
  rename(fips = GEOID, name = NAME)
```

# Questions

## Question 1

```{r ex-01}
fips_prop <- fips_pop %>% 
  pivot_longer(cols = white:hispanic, names_to = "race",
               values_to = "racepop") %>% 
  group_by(name) %>% 
  summarize(race, prop = (racepop/pop)*100000) %>% 
  pivot_wider(names_from = race, values_from = prop)
fips_prop
```

## Question 2

```{r ex-02}
full <- left_join(fips_map, fips_prop) %>% 
  shift_geometry()
```

We know it works because our output is a data frame with the exact same number 
of observations but one more row, and looking at it we can see that the one 
additional row is geometry.

## Question 3

```{r ex-03}
race_map <- function(race){ggplot(full, aes(fill = race, color = race)) +
  geom_sf() +
  theme_bw() +
  scale_fill_viridis(trans = "reverse", option = "B",
                     breaks = c(25000, 50000, 75000, 100000),
                     labels = c("25K", "50K", "75K", "100K")) +
  scale_color_viridis(trans = "reverse", option = "B") +
  theme(legend.position = "bottom") +
  guides(color = "none")
}

race_map(full$white) +
  labs(title = "Density of White People in America",
       fill = "White People per 100,000")

race_map(full$black) +
  labs(title = "Density of Black People in America",
       fill = "Black People per 100,000")
```

## Question 4

```{r ex-04}
binned_race_map <- function(race){full %>% 
  count(cut = cut_interval(race, 5)) %>% 
    ggplot(aes(fill = factor(cut), color = factor(cut))) +
      geom_sf() +
      theme_bw() +
      scale_fill_viridis(discrete = T, option = "B", direction = -1) +
      scale_color_viridis(discrete = T, option = "B", direction = -1) +
      guides(color = "none") +
      theme(legend.position = "bottom",
            legend.text = element_text(size = 6.5))
}

binned_race_map(full$white) +
  labs(title = "Density of White People in America",
       fill = "White People per 100,000")

binned_race_map(full$black) +
  labs(title = "Density of Black People in America",
       fill = "Black People per 100,000")
```

I generally consider this to be a more informative map - since the human eye is 
incapable of properly integrating across the entire continuous scale, even 
though this has less actual data on it it remains a more informative, 
big-picture map.

## Question 5

```{r ex-05}
fips_map %>% 
  left_join(fips_inc) %>% 
  shift_geometry() %>% 
  ggplot(aes(fill = estimate, color = estimate)) +
  geom_sf() +
  theme_bw() +
  scale_fill_viridis(trans = "reverse", option = "A") +
  scale_color_viridis(trans = "reverse", option = "A") +
  theme(legend.position = "bottom") +
  labs(title = "Median Household Income in America",
       fill = "Med Household Income") +
  guides(color = "none")
```

## Question 6

```{r ex-06}
fips_inc %>% 
  arrange(desc(estimate)) %>% 
  slice(1:10)
fips_inc %>% 
  arrange(estimate) %>% 
  slice(1:5)
```