---
title: "king_dav_ps03"
author: "Dav King"
date: "2/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(socviz)
library(slider)
library(covdata)
library(cowplot)
```

# Section A

## 1
```{r A1}
stmf %>% 
  filter(cname == "Belgium") %>% 
  filter(year != 2021) %>% 
  mutate(Year = if_else(year == 2020, "2020", "2000-2019")) %>% 
  group_by(year, week) %>% 
  mutate(deaths = sum(deaths_total)) %>% 
  ggplot(aes(x = week, y = deaths, fill = factor(year), color = Year)) +
  geom_line(size = 0.5) +
  scale_color_manual(values=c("grey80", "red")) +
  labs(title = "Weekly recorded deaths in Belgium, 2000-2020", 
       x = "Week of the Year", y = "Total Deaths") +
  theme_minimal() +
  theme(legend.position = "top", axis.text.x=element_blank(),
        axis.text.y=element_blank())
```

## 2
```{r A2}
a <- stmf %>% 
  filter(cname == "Estonia" & year == "2020" & sex != "b") %>% 
  ggplot(aes(x = week, y = rate_total, color = sex)) +
  geom_line(size = 1) +
  labs(title = "2020 Mortality Rates in Estonia",
       x = "Date", y = "Death Rate", color = "Sex") +
  theme_bw()

b <- stmf %>% 
  filter(cname == "Estonia" & year == "2020" & sex != "b") %>% 
  group_by(sex) %>% 
  summarize(cases = sum(rate_total)/(52*5)) %>% 
  ggplot(aes(x = sex, y = cases, fill = sex)) +
  geom_col() +
  labs(title = "Mortality Rate", x = "Sex", y = "Deaths") +
  guides(fill = "none") +
  theme_bw()

plot_grid(a, b, rel_widths = c(2, 1))
```

While there are some fluctuations in the mortality rates of Estonian men and 
women across 2020, ultimately their mortality rates were almost identical.

## 3
```{r A3}
c <- stmf %>% 
  filter(cname == "Estonia" & year == "2020" &
           age_group == "75-84" & sex != "b") %>% 
  group_by(sex) %>% 
  summarize(dRate = sum(death_rate)/52) %>% 
  ggplot(aes(x = sex, y = dRate, fill = sex)) +
  geom_col() +
  guides(fill = "none") +
  labs(title = "Mortality Rate",
       x = "Sex", y = "Death Rate") +
  theme_bw()

d <- stmf %>% 
  filter(cname == "Estonia" & year == "2020" &
           age_group == "75-84" & sex != "b") %>% 
  ggplot(aes(x = week, y = death_rate, color = sex)) +
  geom_line(size = 1) +
  labs(title = "2020 Mortality Rate", subtitle = "75-84 year olds in Estonia", 
       x = "Week", y = "Death Rate", color = "Sex") +
  theme_bw()

plot_grid(d, c, rel_widths = c(2, 1))
```

Here, there is a clear difference - older men died at much higher rates than 
older women (nearly 2x as many!)

# Section B

```{r intro}
nytcovstate

nytcovstate <- nytcovstate %>% 
  group_by(state) %>% 
  mutate(daily_cases = cases - lag(cases, order_by = date), 
         daily_deaths = deaths - lag(deaths, order_by = date)) 

state_pops <- uspop %>% 
  filter(sex_id == "totsex", hisp_id == "tothisp") %>% 
  select(state_abbr, statefips, pop, state) %>% 
  rename(name = state, state = state_abbr, fips = statefips) %>% 
  mutate(state = replace(state, fips == "11", "DC"))
```

## 1
```{r B1}
nytcovstate <- nytcovstate %>% 
  left_join(state_pops, by = "fips")
```

## 2
```{r B2}
nytcovstate %>% 
  group_by(state.x) %>% 
  summarize(nCase = sum(daily_cases, na.rm = T),
            nDeath = sum(daily_deaths, na.rm = T), 
            popul = sum(pop, na.rm = T)) %>% 
  summarize(state.x, case_rate = (nCase/popul) * 100000,
            death_rate = (nDeath/popul) * 100000)
```

## 3
```{r B3}
nytcovstate %>% 
  filter(state.y == "MD" | state.y == "TX" | state.y == "CA") %>% 
  ggplot(aes(x = date, y = daily_cases, color = state.x)) +
  geom_line() +
  facet_wrap(~ state.x, ncol = 1, scales = "free_y") +
  labs(title = "Daily Case Counts", x = "Date", y = "Daily Cases") +
  guides(color = "none") +
  theme_bw()
```

## 4
```{r B4}
nytcovstate %>% 
  filter(state.y == "MD" | state.y == "TX" | state.y == "CA") %>% 
  group_by(state.x) %>% 
  mutate(avg = slide_dbl(daily_cases, mean, .before = 6)) %>% 
  ggplot(aes(x = date, color = state.x)) +
  geom_line(aes(y = daily_cases), alpha = 0.5) +
  geom_line(aes(y = avg), color = "black", size = 1) +
  facet_wrap(~ state.x, ncol = 1, scales = "free_y") +
  labs(title = "Daily Case Counts", x = "Date", y = "Daily Cases") +
  guides(color = "none") +
  theme_bw()
```

# Section C

## 1
```{r C1}
apple_mobility %>% 
  filter(region == "Montgomery County" & sub_region == "Maryland")
```

## 2
```{r C2}
apple_mobility %>% 
  filter(region == "Montgomery County" & sub_region == "Maryland") %>% 
  ggplot(aes(x = date, y = score, color = transportation_type)) +
  geom_line() +
  facet_wrap(~ transportation_type, ncol = 1, scales = "free_y") +
  labs(title = "Trends in Types of Transit Taken", x = "Date",
       y = "Score Relative to Day 1") +
  guides(color = "none") +
  theme_bw()
```

## 3
This pattern really follows (inversely) the trend in Covid cases over time, such
that all types of transit dropped heavily in March 2020 as Covid hit, but that 
walking and driving (more isolated, personal forms of transit) recovered much 
more quickly than public transit did. On the most basic level, it tells us about
quite explicitly what forms of transit people were taken. If we wanted to 
compare two or more places (see below), we would likely see similar trends - 
though they would depend on the makeup of each location (a county that does not 
already have much public transit, for example, would likely not have much to 
lose; a country which had Covid spikes at different times would show slightly 
different trends).

```{r c3}
apple_mobility %>% 
  filter(sub_region == "Maryland" & region == "Montgomery County" |
         sub_region == "North Carolina" & region == "Durham County") %>% 
  ggplot(aes(x = date, y = score, color = transportation_type)) +
  geom_line() +
  facet_grid(transportation_type ~ region, scales = "free_y") +
  labs(title = "Trends in Types of Transit Taken",
       subtitle = "Durham County, NC vs Montgomery County, MD",
       x = "Date", y = "Score Relative to Day 1") +
  guides(color = "none") +
  theme_bw()
```