---
title: "king_dav_vsd_pset_01"
author: "Dav King"
date: "1/29/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(gapminder)
library(palmerpenguins)
library(socviz)
```

```{r make_plot}
p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))
```

```{r fig1}
p + geom_point() +
  geom_smooth() +
  labs(title = "GDP and Life Expectancy", x = "GDP Per Capita",
       y = "Life Expectancy", caption = "Data: Gapminder")
```

```{r fig2}
p + geom_point() +
  geom_smooth() +
  scale_x_log10(labels = scales::dollar_format()) +
  labs(title = "GDP and Life Expectancy",
       x = "GDP Per Capita (log scale)", y = "Life Expectancy",
       caption = "Data: Gapminder")
```

```{r fig3}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, size = pop)) + geom_point() +
  geom_smooth(method = lm) +
  scale_x_log10(labels = scales::dollar_format()) +
  labs(title = "GDP, Life Expectancy, and Population",
       x = "GDP Per Capita (log scale)", y = "Life Expectancy",
       size = "Population", caption = "Data: Gapminder")
```

```{r fig4}
ggplot(gapminder, aes(x = lifeExp, y = gdpPercap, size = pop)) + 
  geom_point(alpha = 0.2) +
  geom_smooth() +
  scale_y_log10(labels = scales::dollar_format()) +
  labs(title = "GDP, Life Expectancy, and Population",
       x = "Life Expectancy", y = "GDP Per Capita (log scale)",
       size = "Population")
```

# Discussion

## 1
These plots draw from almost the same `ggplot(aes())` mappings. The first two 
use data `p`, which maps `gdpPercap` to x and `lifeExp` to y. The third uses 
these same two mappings, as well as mapping `pop` to size. The fourth does these
same mappings, except the variables mapped to x and y are flipped.

## 2
To varying degrees, these plots do not meet Tufte's principles. To consider each
in turn:

The representation of numbers should be directly proportional to the numerical 
quantities represented: while I am not directly measuring the area of the 
circles mapped to size in figures 3 and 4, given that they are a shape scaled 
radially by a variable, they likely are scaled in terms of area 
disproportionately to the numerical quantities of the data.

Clear, detailed, and thorough labeling: there is fairly consistent labeling 
across the plots, and it is especially good that plots for which this is 
relevant explicitly state that they are using a log scale, as well as denoting 
the numbers at each point on the scale. However, the scales for size are poor = 
they suggest five discrete values of population, which is not true, and they are
hard to tell apart from one another and colored poorly (size should have been 
mapped within `geom_point()` and not `ggplot()`). Additionally, there are no 
explanations of the data on the graphs themselves, including explanations of 
outliers or trendlines.

Show data variation, not design variation: these graphs are good at that! Scales
are consistent and there is a good ink-to-data ratio in these graphs.

Deflated and standardized units of monetary measurement: while these graphs do 
not explicitly depict time on them, they contain data points at different times 
- arguably even worse for potentially violating this criterion. Fortunately, 
these data are inflation-adjusted and denoted in USD, so they meet Tufte's 
criteria.

No more variable dimensions depicted than data dimensions: once again, we run 
into the same issue with the circles - scaled radially, but displayed areally. 
Area is multiple dimensions, but radius is only one - a clear problem. Other 
than that, there are no issues with dimensions of data in these scatterplots.

Don't quote data out of context: unfortunately, there is no information on any 
of these graphs that these data points are duplicates for the same country in 
different years, which could play a clear role in the underlying data structure.
They are well labeled, with calls to where the data comes from, but there is a 
clear problem with directly explaining what the data are.

## 3
Figure 2 is a good, basic display of what the relationship looks like between
the log of GDP Per Capita and Life Expectancy. Figure 3, however, adds a couple
of layers to the graph. It takes the loess regression line and makes it linear,
which allows us to see the trend in the data much more clearly (even if they are
somewhat better modeled by a loess regression line, as all things technically 
are). It also scales the points by population, which is actually a rather 
ineffective addition of data - without colors or reduced alpha levels or 
anything, it's impossible to make sense of this additional data point.

Figure 4 adds a couple of things that improve this graph by a good deal. It maps
the variables onto the opposite axes from before, which is neither good nor bad 
but certainly tells a different story with the data, and reduces the alpha 
level to allow us to see the points by population better and visualize overlap 
(even if the scaling of the data is still somewhat ineffective). It certainly 
suggests that life expectancy is the predictive variable, and also shows just 
how much data is clustered around a life expectancy of 70-80 (which was 
obscured with a default alpha = 1).

```{r fig5}
ggplot(penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species), alpha = 0.3) +
  facet_grid(island ~ species) +
  geom_smooth(method = lm, se = F, color = "black") +
  labs(title = "Are Body Mass and Flipper Length Related in Penguins?",
       subtitle = "Faceted by Species and Island", x = "Body Mass in Grams",
       y = "Flipper Length in Millimeters",
       caption = "Data: German, Williams, and Fraser (2014)") +
  theme_bw() +
  guides(color = "none")
```

This visualization is interesting because it shows a few different data points. 
First, it shows a fairly clear linear relationship between body mass and flipper
length, even if it is somewhat scattered (i.e. large residuals and not ideal 
r-values). It shows that this relationship is almost exactly the same within 
species on different islands, but not the same across species even on the same 
island. It tells us that Adelie penguins live on all three islands, whereas 
Chinstrap are found only on Dream Island and Gentoo only on Briscoe, and it 
tells us that Gentoo penguins are almost always larger (i.e. heavier and with 
longer flippers) than other species of penguin.